### 2022-03-04 paper/localpsf.tex
editing writing

Let $x$ and $x'$ be points in $\Omega$ that are close to each other, and consider the following three approximations, ordered by increasing level of sophistication:
\begin{align}
	\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
	\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
	\phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right). \label{eq:translate3}
\end{align}
Approximations \eqref{eq:translate1} and \eqref{eq:translate2} are standard. Approximation \eqref{eq:translate3}, which we use, is new. We call operators that satisfy \eqref{eq:translate3} ``locally mean-displacement invariant.'' We illustrate local mean-displacement invariance in Figure \ref{fig:mean_displacement_invariance}.

@@

\section{Method}
\label{sec:method}
In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$, how we choose the sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute the impulse response batches $\phi_{x_i}$ (Section \ref{sec:get_impulse_response}), how we interpolate impulse response to approximate integral kernel entries $\Phi(y,x)$, and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$ after the problem has been discretized with finite elements. We also describe how to (optionally) modify $\mathbf{A}_h$ to force it to be symmetric positive semi-definite. The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.

@@

\subsection{Impulse response moments and support estimation}
\label{eq:mean_and_covariance_estimation}

Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$. Hence, $\impulseresponse_x$ is a scaled probability distribution. Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively. We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a fixed constant. The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$ (see Figure FIG). The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$. However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results. The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.

The ellipsoid $E_x$ is not immediately available. To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$ (recall $\gdim \in \{1,2,3\}$ is the spatial dimension of $\Omega$). The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}. Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$. But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points. Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions. We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.

\begin{theorem}[Impulse response moments]
	\label{thm:vol_mean_cov}
	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
	\begin{equation*}
	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
	\end{equation*}
	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$. Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
	\begin{equation*}
	C(x) := 1, \qquad
	L^i(x) := x^i, \qquad
	Q^{ij}(x) = x^i x^j.
	\end{equation*}
	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$. The following results hold:
	\begin{subequations}
		\label{eq:vol_mean_var}
		\begin{align}
		\spatialvol =& \left(\Aop^T C\right)^* \\
		\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
		\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
		\end{align}
	\end{subequations}
	for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
\end{theorem}

@@

We choose sample points, $x_i$, in batches $\pointbatch_k$. 
%A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other. We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}). The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}). 
We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
%to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch. 
%These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$. The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\Aop$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.

We start with a finite set of candidate points $P$ and build $\pointbatch_k$ incrementally with points selected from $P$. For simplicity of explanation, here $B_k$ and $P$ are mutable sets that we add points to and remove points from. First we initialize $\pointbatch_k$ as an empty set. Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches. That is, $p$ is a maximizer of the following optimization problem:
\begin{equation*}
\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
\end{equation*}
Candidate points for the first batch are chosen randomly from $P$.
Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
\begin{enumerate}
	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \tau V_\text{max}$. Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\tau=10^{-5}$).
\end{enumerate}
If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $V(p)$ is not too small) then we add $p$ to $\pointbatch_k$. Otherwise we discard $p$. This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We check whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.

@@

We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to a Dirac combs. The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
\begin{equation*}
	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / V(x_i).
\end{equation*}
We compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
\begin{equation}
	\label{eq:dirac_comb_H_action}
	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*.
\end{equation}
By linearity, the impulse response batch $\combresponse_k$ may be written as follows:
\begin{equation}
	\label{eq:phi_b}
	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
\end{equation}
Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have (approximately)
\begin{equation}
\label{eq:varphi_eval}
	\impulseresponse_{x_i}(z) = 
	\begin{cases}
		V(x_i)\combresponse_k(z), & z \in E_{x_i} \\
		0, & \text{otherwise}
	\end{cases}
\end{equation}
for all $x_i \in \pointbatch_k$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.

Each point source $\delta_{x_i}$ in the Dirac comb is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude. Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section SEC excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold. Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.

@@

\subsection{Approximate integral kernel entries}

Here we describe how we compute approximate kernel entries $\Phi(y,x)$ for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$. To that end, we define the function
\begin{equation}
\label{eq:fxyxp}
f_{y,x}(x') := \frac{V(x)}{V(x')}\phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
\end{equation}
It follows from inspection that 
\begin{equation*}
\Phi(y,x) = f_{y,x}(x),
\end{equation*}
and from the discussion in Section SEC, we expect that $\Phi(y,x) \approx f_{y,x}(x')$ when $x'$ is close to $x$. The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section SEC, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly. 
%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
%As discussed in Section SECTION, we make the approximation
%\begin{equation*}
%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
%\end{equation*}
%if $x'$ is close to $x$. Hence, w

We approximate $\Phi(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$. The data used in the interpolation are the following (point,value) pairs
\begin{equation}
\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
\end{equation}
and $(x, f_{y,x}(x))$ is the (point,value) pair we approximate via interpolation.
Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
%\begin{equation}
%\label{eq:phi_mu_shift32}
%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
%\end{equation}
%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$. 
%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree [CITE] of all sample points.

From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$. But it may happen that $y - \mu(x) + \mu(x_i) \notin \Omega$, even if $y,x,x_i \in \Omega$. In this case, the point $x_i$ is considered invalid, and not used in the interpolation, even if it is close to $x$. We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) [CITE] of the mesh cells used to discretize the problem.

>>>>

Let $x$ and $x'$ be points in $\Omega$ that are close to each other, and consider the following three approximations, ordered by increasing level of sophistication:
\begin{align}
	\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
	\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
	\phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right). \label{eq:translate3}
\end{align}
We use approximation \eqref{eq:translate3}, which we call ``locally mean-displacement invariance.'' This is illustrated in Figure \ref{fig:mean_displacement_invariance}.

@@

\section{Method}
\label{sec:method}
In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$. We also describe how to (optionally) modify $\mathbf{A}_h$ to force it to be symmetric positive semi-definite. The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.

@@

\subsection{Impulse response support ellipsoids}
\label{eq:mean_and_covariance_estimation}

%Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$. Hence, $\impulseresponse_x$ is a scaled probability distribution. Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.

We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a fixed constant. The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$.

The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$. However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results. The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.

\section{Impulse response moments}

%The ellipsoid $E_x$ is not immediately available. 
To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$.
%(recall $\gdim$ is the spatial dimension of $\Omega$).
The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}. Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$. But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$ that we wish to determine $E_x$ at, and we will need to determine $E_x$ at a large number of points. Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions. We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.

\begin{theorem}[Impulse response moments]
	\label{thm:vol_mean_cov}
%	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
%	\begin{equation*}
%	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
%	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
%	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
%	\end{equation*}
%	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	
	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$. Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
	\begin{equation*}
	C(x) := 1, \qquad
	L^i(x) := x^i, \qquad
	Q^{ij}(x) = x^i x^j.
	\end{equation*}
	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$. The following results hold:
	\begin{subequations}
		\label{eq:vol_mean_var}
		\begin{align}
		\spatialvol =& \left(\Aop^T C\right)^* \\
		\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
		\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
		\end{align}
	\end{subequations}
	for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
\end{theorem}

@@

We choose sample points, $x_i$, in batches $\pointbatch_k$. 
%A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other. We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}). The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}). 
We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points within a batch.
%to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch. 
%These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$. The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\Aop$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.

We start with a finite set of candidate points $P$ and build $\pointbatch_k$ incrementally with points selected from $P$. For simplicity of explanation, here $\pointbatch_k$ and $P$ are mutable sets that we add points to and remove points from. First we initialize $\pointbatch_k$ as an empty set. Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$. 
%That is, $p$ is a maximizer of the following optimization problem:
%\begin{equation*}
%\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
%\end{equation*}
Candidate points for the first batch $S_1$ are chosen randomly from $P$.
Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
\begin{enumerate}
	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \tau V_\text{max}$. Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\tau=10^{-5}$).
\end{enumerate}
If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $V(p)$ is not too small) then we add $p$ to $\pointbatch_k$. Otherwise we discard $p$. This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We check whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.

@@

We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs. The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
\begin{equation*}
	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / V(x_i).
\end{equation*}
We compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
\begin{equation}
	\label{eq:dirac_comb_H_action}
	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
\end{equation}
The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}.
%\begin{equation}
%\label{eq:phi_b}
%\combresponse_k = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
%\end{equation}
%\begin{equation}
%	\label{eq:phi_b}
%	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
%\end{equation}
Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
\begin{equation}
\label{eq:varphi_eval}
	\impulseresponse_{x_i}(z) / V(x_i) = 
	\begin{cases}
		\combresponse_k(z), & z \in E_{x_i} \\
		0, & \text{otherwise}
	\end{cases}
\end{equation}
for all $x_i \in \pointbatch_k$. By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.

Each point source $\delta_{x_i}$ in the Dirac comb is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude. Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold. Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.

@@

\subsection{Approximate integral kernel entries}
\label{sec:approximate_kernel_entries}

In Section \ref{sec:intro} we noted that the integral kernel, $\Aker$, is inaccessible. Here we describe how we compute approximate kernel entries,
\begin{equation*}
\widetilde{\Phi}(y,x) \approx \Phi(y,x),
\end{equation*}
for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
% while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
To that end, we define the function
\begin{equation}
\label{eq:fxyxp}
f_{y,x}(x') := \frac{V(x)}{V(x')}\phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
\end{equation}
It follows from inspection that 
\begin{equation*}
\Phi(y,x) = f_{y,x}(x),
\end{equation*}
and from the discussion in Section SEC, we expect that $\Phi(y,x) \approx f_{y,x}(x')$ when $x'$ is close to $x$. The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section SEC, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly. 
%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
%As discussed in Section SECTION, we make the approximation
%\begin{equation*}
%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
%\end{equation*}
%if $x'$ is close to $x$. Hence, w

We evaluate the approximate kernel entry $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$. The data used in the interpolation are the following (point,value) pairs
\begin{equation}
\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
\end{equation}
and $(x, f_{y,x}(x)) = (x,\Phi(y,x))$ is the (point,value) pair we approximate via interpolation.
Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
%\begin{equation}
%\label{eq:phi_mu_shift32}
%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
%\end{equation}
%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$. 
%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points.

From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$. But it may happen that $y - \mu(x) + \mu(x_i) \notin \Omega$, even if $y,x,x_i \in \Omega$. In this case, the point $x_i$ is considered invalid, and not used in the interpolation, even if it is close to $x$. We check whether $x_i$ is valid (i.e., whether $y - \mu(x) + \mu(x_i) \in \Omega$) by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
